---
title: "Assignment 01"
author: "Cameron Atkins"
date: "Due: 2023-09-08"
output: html_document
---

### 1.Data used for the assignment
The data used for this project is IMDb's list of Top 1000 movies of all time. The
original data featured nine colums of data, but for the purposes of this assignment,
columns of data was reduced to four.

[Link to Kaggle DataSet](https://www.kaggle.com/datasets/inductiveanks/top-1000-imdb-movies-dataset)

#### Importing the data from the working directory

```{r}
data <- read.csv("Top_1000_IMDb_movies_New_version.csv", header = TRUE)
```

#### Taking a look at the data
```{r}
head(data)
```

#### Select 4 Variables: Year of Release, Watch Time, MetaScore, and Gross Ticket Sales
```{r}
keep_columns <- c(3, 4, 6, 7)
```

Reduce the number of variables used
```{r}
data <- data[, keep_columns]
```

View new data after reduction

```{r}
head(data)
```
### 2. Summary of the data
#### Structure of the data
```{r}
str(data)
```

#### Summary of the data
```{r}
summary(data)
```

### 3. Search for any outliers in the box office gross
```{r}
# Convert gross figures from character to numeric
numeric_gross <- as.numeric(data$Gross)
xu= quantile(numeric_gross, probs = 0.85, na.rm = TRUE)
xl = quantile(numeric_gross, probs = 0.15, na.rm = TRUE)
IQR = xu - xl

plot(numeric_gross, ylab='Box Office Gross')
abline(h=xu, lty=2)
abline(h=xu+1.5*IQR, lty=3)
```

Out of the 1000 movies, only 13 earned more than US $400 million. However, dataset
doesn't specify if amounts are in current or real dollars.

### 4. Visualizations for some variables

#### Scatterplot between 2 variables
A visualization that can help determine if films produced recently have higher 
metascores than older films.
```{r}
plot(data$Year.of.Release, data$Metascore.of.movie,
     main = "Years and MetaScore",
     xlab = "Year of Release",
     ylab = "MetaScore",
     pch = 16,
     col = "deepskyblue"
     )
```

#### One barplot providing the count of the factor variable

```{r}
year_factor <- factor(c(data$Year.of.Release))
year_table <- table(year_factor)
year_table
barplot(year_table,
        main = "Barplot of a Factor",
        xlab = "Categories",
        ylab = "Frequency",
        col = "deepskyblue",  # Set the bar color (you can customize)
        border = "black"  # Set the bar border color (you can customize)
)
```

#### Histogram of the length of the Top 1000 movies

```{r}
hist(data$Watch.Time,
     breaks = seq(45,350, by = 20),
     main = "Length of Top 1000 Movies",
     xlab = "Movie Length (in Minutes)",
     ylab = "Frequency",
     col = "deepskyblue",  # Set the bar color (you can customize)
     border = "black"    # Set the bar border color (you can customize)
)
```

### 5. Create a linear model for two of the numerical variables

Plotting watch time as x and metascore as y to determine if there is any 
correlation between the length of the movie and its Metascore.
```{r}
df <- data.frame(x = c(data$Watch.Time), y=c(data$Metascore.of.movie))
model <- lm(y ~ x, data = df)
summary(model)
plot(main = "Relationship between Movie Watch Time and MetaScore", 
     df$x, 
     df$y, 
     xlab = "Movie Watch Time (minutes)", 
     ylab = "Movie MetaScore")
abline(model, col = "red")
```

